---
title: "Analysis BioRodDis data"
author: "Vincent Sluydts"
institute: "University of Antwerp"
date: "`r Sys.Date()`"
output: html_document
---


```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(Hmsc) #depends on package coda!
library(corrplot)
library(colorspace)
library(vioplot)
library(GGally)
library(writexl)
library(ggplot2)
library(PerformanceAnalytics)
library(grid)
library(gridExtra)

```


---
# INPUT DATA

Always good practice to start with an empty, clean work space:

```{r}
rm(list = ls()); gc()
```

Read in the data files needed for the analysis
```{r}
#read in  the bioroddis data file (rds format) and the information on pathogen associated traits (csv)
dat<-readRDS("./data/bioroddis.rds")
#dat<-readRDS("./data/all11.rds")
trait <- read.csv(file="./data/disease_traits.csv", sep=",", row.names = 1)

```


## pathogens names 

Here we combine the different strains for the analysis to have genus level pathogen info per specimen

```{r}
dat$Borrelia <- ifelse((dat$Borrelia_1+dat$Borrelia_2) > 0, 1,0)  
dat$Mycoplasma_ravipulmonis <- ifelse((dat$Mycoplasma_ravipulmonis_1 + dat$Mycoplasma_ravipulmonis_2) >0, 1, 0)
dat$Chlamydia <- ifelse((dat$Chlamydia_1+dat$Chlamydia_2) > 0, 1,0)
dat$Rickettsia <- ifelse((dat$Rickettsia_1+dat$Rickettsia_2) > 0, 1,0)
dat$Streptobacillus <- ifelse((dat$Streptobacillus_1 + dat$Streptobacillus_2)> 0, 1,0)
dat$Hanta <- ifelse((dat$IFA_PUUV +dat$IFA_DOBRV)> 0, 1,0)
dat$Lepto <- ifelse((dat$lepto.lipl32 + dat$Leptospira)>0, 1, 0)
```

```{r}
names(dat)
```

```{r}
disease.names <- c("Bartonella","Mycoplasma_haemo","Mycoplasma_coco",  "Candidatus_Neoehrlichia",   "Anaplasma"    , "Francisella", "Borrelia"  ,  "Orientia" ,   "Sarcocystidae" , "Mycoplasma_penetrans", "Mycoplasma_ravipulmonis" , "Mycoplasma_microti", "Mycoplasma_insons"  , "Ehrlichia" ,   "Chlamydia"     ,  "Rickettsia"   ,  "Leptospira"  ,   "Streptobacillus",  "IFA_POX" ,"Hanta" , "IFA_LCMV" , "Lepto")  

#extract columns with pathogens of interest screened
diseases <- dat[names(dat) %in% disease.names]
```


## Overal pathogen prevalence

```{r}
cbind(colSums(diseases, na.rm=T), round((colSums(diseases, na.rm=T)/nrow(diseases))*100, 2)) -> dis.prev
dis.prev <- as.data.frame(dis.prev)
colnames(dis.prev)<- c("positive", "prevalence")
dis.prev[order(dis.prev$positive),]

```

Pathogens with over 50 positive occurrences are considered for further analysis

```{r}
dis.prev[dis.prev$positive>50,]
```


Host species diversity per country


```{r}

names(dat)

cbind(dat %>% filter(Country == 'Belgium') %>% dplyr::select(Species) %>% table(),
dat %>% filter(Country == 'France') %>% dplyr::select(Species) %>% table(),
dat %>% filter(Country == 'Germany') %>% dplyr::select(Species) %>% table(),
dat %>% filter(Country == 'Ireland') %>% dplyr::select(Species) %>% table(),
dat %>% filter(Country == 'Poland') %>% dplyr::select(Species) %>% table()) -> soorten

soort <- as.data.frame(soorten)
colnames(soort) <- c("Belgium","France", "Germany", "Ireland", "Poland")
soort
```


Setting the stage for the HMSC analysis

```{r}
#dat %>% filter(Country != 'Poland') -> dat2 #as there was only one habitat sampled and not all samples from Poland were tested for all diseases, Poland is left out of the HMSC analysis.
dat %>% filter((Genus == "Apodemus" | Genus == "Myodes" ) & SpatTemp != "BEPFO6_Winter_2020" & Site_Code != "GEFKAT" & Site_Code != "GEFKIR"& Country != 'Poland') %>% droplevels()-> dat2

#Here we omit the data from Poland as we did not have IFA and Leptospirosis tested, as well as fieldwork in one locality [habitat], season/year only.
#Two German sites, GEFKAT and GEFKIR were omitted for sampling only once as well, after it was decided to move to sites where rodents were more abundant. Only 6 rodents were trapped in those sites. 
#BEPFO6 is an additional sampling behind a bird cage at university. Two rats were trapped and omitted from the HMSC analysis as well.
# droplevels() ensures no residual factor levels remain as this will generate error during model runs



patho<-c("Bartonella","Mycoplasma_haemo", "Mycoplasma_coco","IFA_POX","Sarcocystidae","Candidatus_Neoehrlichia" ,"Hanta", "Borrelia","Lepto", "Anaplasma" ,"Francisella") #pathogens with over 50 occurences

host <- c("Sex", "Weight","Maturity", "Richness_gut_species", "lrich") 

envir <- c("grass","f.broad", "f.coni", "w.perm","w.temp", "hab2", "hab3", "HSD" )
clim <- c("total_rain","max_temp1",  "min_temp1")
tijd<- c("Season", "Year")
anthro <- c("imper", "pop", "HFI")
covariates <- c(host, envir, clim, anthro, tijd)


randomeffects<-c("Id_code","Species","Genus" ,"Site_Code", "Period", "SpatTemp", "Country")


needed<-c(covariates, randomeffects, patho)
dat2 %>% dplyr::select(all_of(needed)) -> dat3
head(dat3)

```



Check for NA values:

```{r}
dat3 %>%
  summarise_all(funs(sum(is.na(.))))
```
```{r}
dat3 %>% dplyr::select(all_of(needed)) %>% na.omit()-> dat4
```

Final check 
```{r}
dat4 %>% select(Species) %>% table()
dat4 %>% select(Site_Code) %>% table()
dat4 %>% select(SpatTemp) %>% table()

dat4 %>% select(Season) %>% table()
dat4 %>% select(Year) %>% table()
dat4 %>% select(Period) %>% table()

any(is.na(dat4))
```
# HMSC ANALYSIS -----

---
### STEP 1: Setup of Y and X matrices

```{r}
#Y matrix of HMSC diseases
Y<-as.matrix(dat4[,names(dat4)%in%patho])
colnames(Y)
```

--
```{r}

#COVARIATE STRUCTURE 

host <- c("Sex", "Weight","Maturity", "Richness_gut_species", "lrich")
envir <- c("grass","f.broad", "f.coni", "w.perm","w.temp", "hab2", "hab3", "HSD")
clim <- c("total_rain","max_temp1",  "min_temp1")
tijd<- c("Season", "Year")
anthro <- c("imper", "pop", "HFI")
covariates <- c(host, envir, clim, anthro, tijd)

randomeffects<-c("Id_code","Genus","Species","Site_Code", "Period", "SpatTemp", "Country")

X<-dat4[, names(dat4)%in%c(covariates, randomeffects)]
colnames(X)
```

```{r}
X %>% select('Id_code', 'Site_Code', 'Country', 'Sex', 'Weight', 'grass', 'max_temp1') %>% head()
```


---
### Check Species/Pathogen Matrix:

```{r}
colnames(Y)
# Check for absent (0) or ubiquitous species/pathogens (1).

# Note that we have selected on those pathogens for which at least 50 occurrences were observed.

colSums(Y)
```

---
### STEP 2: Set the HMSC model structure


#### Read in traits  
```{r }

TR<-trait

TR$Detection <- as.factor(TR$Detection) #detection method: 16S, IFA or PCR
TR$Transmission <- as.factor(TR$Transmission) #vector, environmental or direct as dominant transmission pathways
TR$patho <- as.factor(TR$patho) #virus, bacteria, protozoa
TR$domHost <- as.factor(TR$domHost) #dominant host, lously defined as the host in which the pathogen occured with highest prevalence


TR <- TR[match(colnames(Y), rownames(TR)), ]
TR

```


#### Fixed effects model structure
```{r }
## MODEL  ------
names(X)

XFormula0 = ~ hab2 #HABITAT

XFormula1 = ~ Species + Sex + Weight + Maturity + lrich  #SPECIES + HOST CHARACTERISTICS [INT]

XFormula2 = ~ grass + f.broad + f.coni + w.perm + w.temp + total_rain + min_temp1 + max_temp1 #ENVIRONMENTAL VARS + CLIMATE [EXT]

XFormula3 = ~  HFI +  HSD + HFI:HSD #ANTHROPOGENIC FACTORS [ANT]

XFormula4 = ~ Species + Sex + Weight + Maturity + lrich +  grass + f.broad + f.coni + w.perm + w.temp + total_rain + min_temp1 + max_temp1 + HFI + HSD #INT+EXT+ANTHRO

```

--
#### Trait effect model structure

```{r}
#TrForm1    = ~ Transmission
#for plotting perhaps easiest to remove intercept here
TrForm1    = ~ -1 + Transmission
```


--
#### Random effects model structure
```{r }

#Define the study design
studyDesign = data.frame( Spat = dat4$Site_Code, Temp = dat4$Period)

# Define Random level at time unit [season*year]
rL.time = HmscRandomLevel(units = levels(studyDesign$Temp))
# Define Random level at trap site unit
rL.spatial = HmscRandomLevel(units = levels(studyDesign$Spat))


```

### Different models defined as in the table 1 from the paper

```{r}

BR0 = Hmsc(Y = Y, XData = X, XFormula = XFormula0, distr="probit", TrData = TR, TrFormula = TrForm1, studyDesign = studyDesign, ranLevels = list(Spat = rL.spatial,  Temp = rL.time))

BR1 = Hmsc(Y = Y, XData = X, XFormula = XFormula1, distr="probit",  TrData = TR, TrFormula = TrForm1, studyDesign = studyDesign, ranLevels = list(Spat = rL.spatial,  Temp = rL.time))

BR2 = Hmsc(Y = Y, XData = X, XFormula = XFormula2, distr="probit", TrData = TR, TrFormula = TrForm1, studyDesign = studyDesign, ranLevels = list(Spat = rL.spatial,  Temp = rL.time))

BR3 = Hmsc(Y = Y, XData = X, XFormula = XFormula3, distr="probit",  studyDesign = studyDesign,TrData = TR, TrFormula = TrForm1, ranLevels = list(Spat = rL.spatial,  Temp = rL.time))

BR4 = Hmsc(Y = Y, XData = X, XFormula = XFormula4, distr="probit",TrData = TR, TrFormula = TrForm1,  studyDesign = studyDesign, ranLevels = list(Spat = rL.spatial,  Temp = rL.time))

```


---
## STEP 3: TESTRUN 

```{r include=FALSE}

#### TEST RUN
nChains = 1
test.run = TRUE
if (test.run){
#with this option, the vignette runs fast but results are not reliable
thin = 1
samples = 10
transient = 5
verbose = 0
} else {
#with this option, the vignette evaluates slow but it reproduces the results of the
#pdf version
thin = 10
samples = 1000
transient = 500*thin
verbose = 0
}
```


--
#### Run MCMC samples

```{r include = F, eval=T}

EU0.test = sampleMcmc(BR0, thin = thin, samples = samples, transient = transient,
               nChains = nChains)

EU1.test = sampleMcmc(BR1, thin = thin, samples = samples, transient = transient,
               nChains = nChains)

EU2.test = sampleMcmc(BR2, thin = thin, samples = samples, transient = transient,
               nChains = nChains)

EU3.test = sampleMcmc(BR3, thin = thin, samples = samples, transient = transient,
               nChains = nChains)

EU4.test = sampleMcmc(BR4, thin = thin, samples = samples, transient = transient,
               nChains = nChains)

```

---
## STEP 4: SAVE INTO object for automatic running 

Here we save all the information for running the models and this will be loaded in a script which can be run as a background job and uses different cores for each chain [see script: HMSC_run_models.R].

```{r }
# Place the models in a list together
models = list(BR0, BR1, BR2, BR3, BR4)
modelnames = c("HAB", "INT", "EXT", "ANT", "INT-EXT-ANT") # Give them the proper names
```

```{r }
# Save the models in the model folder
save(models, modelnames,file = paste('./models', 'unfitted_ALL', sep="/"))
```

After this object is read in using the script HMSC_run_models.R for sampling the chains in parallel over longer periods. Explanatory and predictive performance follow later on using the script HMSC_compute_model_fit.R. All these scripts follow the design proposed by the authors of the HMSC package. for more info check: 

https://www.helsinki.fi/en/researchgroups/statistical-ecology/software/hmsc


### SPLIT FOR CONVENIENCE DUE TO long runtimes

Basically here we split the saved object and list one or two models, this makes it easier to run each one next to each other in the background.

```{r }
# Place the models in a list together
models = list(BR0, BR1)
modelnames = c("HAB", "INT") # Give them the proper names
```

```{r }
# Save the models in the model folder
save(models, modelnames,file = paste('./models', 'unfitted_BR01', sep="/"))
```

```{r }
# Place the models in a list together
models = list(BR2, BR3)
modelnames = c("EXT", "ANT") # Give them the proper names
```

```{r }
# Save the models in the model folder
save(models, modelnames,file = paste('./models', 'unfitted_BR23', sep="/"))
```



```{r }
# Place the models in a list together
models = list(BR4)
modelnames = c("INT-EXT-ANT") # Give them the proper names
```

```{r }
# Save the models in the model folder
save(models, modelnames,file = paste('./models', 'unfitted_BR4', sep="/"))
```
